---
title: The Human Demographic Map
---


```{r}
    knitr::opts_chunk$set(message=FALSE, warning=FALSE, results="hide") 
```

This figure is based on Livi-Bacci's, but updated to use the best data.

The essential idea is that, under a broad range of assumptions, population growth is determined once you know

We begin with data from Max Roser's [Our World in Data](https://ourworldindata.org), which has downloadable files. We can get national life expectancy, total fertility and population growth in seperate sets.

```{r}

source('./R/project_functions.r')
source('./R/custom_palettes.r')

download_file("https://ourworldindata.org/grapher/children-per-woman-UN.csv", "./csv/children-per-woman-UN.csv")

download_file("https://ourworldindata.org/grapher/total-fertility-rate.csv", "./csv/total-fertility-rate.csv")

download_file("https://ourworldindata.org/grapher/life-expectancy.csv", "./csv/life-expectancy.csv")

download_file("https://ourworldindata.org/grapher/population-by-country.csv", "./csv/population-by-country.csv")

download_file("https://ourworldindata.org/grapher/natural-population-growth.csv", "./csv/natural-population-growth.csv")

```

In each case the coding is similar: national units are under the variable `Entity` with a corresponding `Country.code`, and we have the outcome variable. I will use the codes `le` for life expectancy, `r` for growth and `


```{r}

life <- read.csv("./csv/life-expectancy.csv", stringsAsFactors=FALSE)

# Entity
# Year
# Country.code
# Life.Expectancy.at.Birth..both.genders.

life$le <- life$Life.Expectancy.at.Birth..both.genders.
life$key <- paste(life$Country.code, life$Year)

growth <- read.csv("./csv/natural-population-growth.csv", stringsAsFactors=FALSE)

# Entity
# Year
# Country.code
# Rate.of.Natural.Population.Increase...UN.2015

growth$gr <- growth$Rate.of.Natural.Population.Increase...UN.2015
growth$key <- paste(growth$Country.code, growth$Year)

tfr <- read.csv("./csv/total-fertility-rate.csv", stringsAsFactors=FALSE)

tfr$tfr <- tfr$Gapminder..Fertility.v6.
tfr$key <- paste(tfr$Country.code, tfr$Year)

# Entity
# Year
# Country.code
# Gapminder..Fertility.v6.

```

We can combine all of these using the keys.


```{r}

d <- life
d$tfr <- tfr$tfr[match(d$key, tfr$key)]
d$gr <- growth$gr[match(d$key, growth$key)]

d$my_col <- data_gradient(d$gr, my.start=0, my.stop=3.5)

d$my_col[is.na(d$gr)] <- gray(0.8)

# i cna probabl estimate the relationships between these empirically...then recalculate!

```

TFR, LE and growth are related by a log-linear equation

$$
	growth = a + b log(tfr) + c log(le)
$$

Empirical estimation puts this at


```{r}

d$ln_tfr <- log(d$tfr)
d$ln_le <- log(d$le)

# together these caluclate r exaactly?

m1 <- lm(d$gr ~ d$ln_tfr + d$ln_le) # pretty close!

```

And, since we are missing growth for many populations, we can include our estimate from the fitted model. I call the function `growth_isocline`.

```{r}

d$x <- coef(m1)[1] + coef(m1)[2]*d$ln_tfr + coef(m1)[3]*d$ln_le

# r = a + b*ln_le + c*ln_tfr
# -c*ln_tfr + r = a + b*ln_le 
# -c*ln_tfr = a + b*ln_le - r

# exp(ln_tfr) = exp(( a + b*ln_le - r)/(-c))


# exp(ln_tfr) = exp((-10.96 + 2.36*ln_le - r)/2.39)

growth_isocline <- function(x, r) exp( (coef(m1)[1] + coef(m1)["d$ln_le"] * log(x) - r) / -coef(m1)["d$ln_tfr"] )

```

The result is, for me, one of the most beautifully informative figures ever.


```{r, fig.height=15/2.54, fig.width=21/2.54}

d$my_col <- data_gradient(d$x, my.start=0, my.stop=3, colors=orangered )

# png('./test.png', height=15, width=21, units="cm", res=500)

pdf('./pdf/test.pdf', height=15/2.54, width=21/2.54)

	tar <- 1:nrow(d)

	plot(jitter(d$le[tar], amount=0.1), jitter(d$tfr[tar], amount=0.1), 
		col=col_alpha(d$my_col[tar], 0.3), 
		pch=16, frame.plot=FALSE, cex=1, ylim=c(1,9), xlim=c(20,80), 
		xlab="Life expectancy at birth", 
		ylab="Number of children per woman", las=1, axes=FALSE)
	# curve(growth_isocline(x, -1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 0), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 2), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 3), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 4), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)


	# abline(v=gurv$e0, col="blue")

	# e0 calculated from published l15 in each case using equation  -14.581 +  79.668*l15, not the best but okay

	# bentley <- c(6.2, 4.2, 4.3, 4.36, 5.5, 5.8, 5.69, 6.1)

	# abline(h=bentley, col="green")

	# Europe
	europe <- c("United Kingdom",
	"France",
	"Finland",
	"Germany",
	"Greece",
	"Sweden",        
	"Norway",
	"Switzerland")

	# commonwealth
	com <- c("United States",
	"Canada",
	"Australia",
	"India")

	# east asia
	asia <- c("Japan",
	"China",
	"Vietnam",
	"South Korea")

	# south america
	s_america <- c("Bolivia", "Brazil")

	tar <- which(d$Year == 2000)
	points(d$le[tar], d$tfr[tar], pch=20)

	# text

	# text(34.65842, 1.709144, label="-1%", srt=(-14.5))
	text(39.70611, 2.407064, label="0% population growth", srt=(-14.5))
	text(43.22775, 3.401599, label="1%", srt=(-14.5))
	text(48.51021, 4.622958, label="2%", srt=(-16))
	text(55.67089, 6.245621, label="3%", srt=(-18))
	text(64.35761, 8.321932, label="4%", srt=(-20))


	axis(2, at=seq(1,9, by=1), las=1)
	axis(1, at=seq(20,90, by=10), las=1)


	text(56.78387, 3.349255, "world nations,\nyear 2000", font=2)

dev.off()

# reduce the number of nations, esp small ones that are probably throwing off
# maybe 


```

We can include other informatoin as well.


Here, I draw these as regions.


```{r, fig.height=15/2.54, fig.width=21/2.54}

d$my_col <- data_gradient(d$x, my.start=0, my.stop=3, colors=orangered )

pdf('./pdf/test2.pdf', height=15/2.54, width=21/2.54)

	tar <- 1:nrow(d)

	plot(d$le, d$tfr, col=NA, frame.plot=FALSE, 
		ylim=c(1,9), xlim=c(20,80), 
		xlab="Life expectancy at birth", 
		ylab="Number of children per woman", las=1, axes=FALSE)

	# curve(growth_isocline(x, -1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)

	foragers <- read.csv("./csv/forager-polygon.csv", stringsAsFactors=FALSE)
	pastoralists <- read.csv("./csv/pastoralist-polygon.csv", stringsAsFactors=FALSE)
	hort <- read.csv("./csv/horticulturalist-polygon.csv", stringsAsFactors=FALSE)

	tar <- 1:nrow(d)
	points(jitter(d$le[tar], amount=0.1), jitter(d$tfr[tar], amount=0.1), pch=16, col=col_alpha("gray", 0.3))


	polygon(hort$x, hort$y, border=NA, col=col_alpha("orange", 0.6) )
	polygon(foragers$x, foragers$y, border=NA, col=col_alpha("blue", 0.6) )
	polygon(pastoralists$x, pastoralists$y, border=NA, col=col_alpha("red", 0.6) )


	tar <- which(d$Year == 2000)
	points(d$le[tar], d$tfr[tar], pch=20)

	curve(growth_isocline(x, 0), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 2), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 3), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 4), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)



	# text

	# text(34.65842, 1.709144, label="-1%", srt=(-14.5))
	text(39.70611, 2.407064, label="0% population growth", srt=(-14.5))
	text(43.22775, 3.401599, label="1%", srt=(-14.5))
	text(48.51021, 4.622958, label="2%", srt=(-16))
	text(55.67089, 6.245621, label="3%", srt=(-18))
	text(64.35761, 8.321932, label="4%", srt=(-20))


	text(26.19916, 2.804078, label="hunter-\n gatherers", col="blue")

	text(30.95021, 8.619962, label="horticulturalists", col="orange")

	axis(2, at=seq(1,9, by=1), las=1)
	axis(1, at=seq(20,90, by=10), las=1)


	text(56.78387, 3.349255, "world nations,\nyear 2000", font=2)

dev.off()

# reduce the number of nations, esp small ones that are probably throwing off
# maybe 


```




```{r, fig.height=15/2.54, fig.width=21/2.54}

d$my_col <- data_gradient(d$x, my.start=0, my.stop=3, colors=orangered )

pdf('./pdf/test4.pdf', height=15/2.54, width=21/2.54)

	tar <- 1:nrow(d)

	plot(d$le, d$tfr, col=NA, frame.plot=FALSE, 
		ylim=c(1,9), xlim=c(20,85), 
		xlab="Life expectancy at birth", 
		ylab="Number of children per woman", las=1, axes=FALSE)

	# curve(growth_isocline(x, -1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)

	foragers <- read.csv("./csv/forager-polygon.csv", stringsAsFactors=FALSE)
	pastoralists <- read.csv("./csv/pastoralist-polygon.csv", stringsAsFactors=FALSE)
	hort <- read.csv("./csv/horticulturalist-polygon.csv", stringsAsFactors=FALSE)

	# tar <- 1:nrow(d)
	# points(jitter(d$le[tar], amount=0.1), jitter(d$tfr[tar], amount=0.1), pch=16, col=col_alpha("gray", 0.3))


	polygon(hort$x, hort$y, border=NA, col=col_alpha(gray(0.8), 0.6) )
	polygon(foragers$x, foragers$y, border=NA, col=col_alpha(gray(0.6), 0.6) )
	polygon(pastoralists$x, pastoralists$y, border=NA, col=col_alpha(gray(0.5), 0.6) )


	curve(growth_isocline(x, 0), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 2), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 3), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 4), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)




	col_1955 <- aquamarine[5]

	tar <- which(d$Year == 1955)
	points(d$le[tar], d$tfr[tar], pch=16, col=col.alpha(col_1955, 0.2))

	tar <- which(d$Year == 1955 & d$Country.code=="USA")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="USA", col=col_1955, pos=2)

	tar <- which(d$Year == 1955 & d$Country.code=="JPN")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="Japan", col=col_1955, pos=2)

	tar <- which(d$Year == 1955 & d$Country.code=="CHN")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="China", col=col_1955, pos=4)

	tar <- which(d$Year == 1955 & d$Country.code=="BOL")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="Boliva", col=col_1955, pos=4)

	tar <- which(d$Year == 1955 & d$Country.code=="TZA")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="Tanzania", col=col_1955, pos=1)

	tar <- which(d$Year == 1955 & d$Country.code=="PER")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="Peru", col=col_1955, pos=1)

	tar <- which(d$Year == 1955 & d$Country.code=="IND")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="India", col=col_1955, pos=1)

	tar <- which(d$Year == 1955 & d$Country.code=="SWE")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="Sweden", col=col_1955, pos=1)

	tar <- which(d$Year == 1955 & d$Country.code=="MAR")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="Morocco", col=col_1955, pos=1)

	tar <- which(d$Year == 1955 & d$Country.code=="NAM")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="Namibia", col=col_1955, pos=1)

	tar <- which(d$Year == 1955 & d$Country.code=="FJI")
	points(d$le[tar], d$tfr[tar])
	text(d$le[tar], d$tfr[tar], label="Fiji", col=col_1955, pos=1)



	col_2005 <- orangered[5]

	tar <- which(d$Year == 2005)
	points(d$le[tar], d$tfr[tar], pch=16, col=col_alpha(col_2005, 0.2))

	tar <- which(d$Year == 2005 & d$Country.code=="USA")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="USA", col=col_2005, pos=3)

	tar <- which(d$Year == 2005 & d$Country.code=="JPN")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="Japan", col=col_2005, pos=1)

	tar <- which(d$Year == 2005 & d$Country.code=="CHN")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="China", col=col_2005, pos=1)

	tar <- which(d$Year == 2005 & d$Country.code=="BOL")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="Boliva", col=col_2005, pos=4)

	tar <- which(d$Year == 2005 & d$Country.code=="TZA")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="Tanzania", col=col_2005, pos=4)

	tar <- which(d$Year == 2005 & d$Country.code=="PER")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="Peru", col=col_2005, pos=3)

	tar <- which(d$Year == 2005 & d$Country.code=="IND")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="India", col=col_2005, pos=2)

	tar <- which(d$Year == 2005 & d$Country.code=="SWE")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="Sweden", col=col_2005, pos=4)

	tar <- which(d$Year == 2005 & d$Country.code=="MAR")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="Morocco", col=col_2005, pos=1)

	tar <- which(d$Year == 2005 & d$Country.code=="NAM")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="Namibia", col=col_2005, pos=1)

	tar <- which(d$Year == 2005 & d$Country.code=="FJI")
	points(d$le[tar], d$tfr[tar], col=col_alpha(col_2005, 0.5))
	text(d$le[tar], d$tfr[tar], label="Fiji", col=col_2005, pos=1)




	# text

	# text(34.65842, 1.709144, label="-1%", srt=(-14.5))
	text(39.70611, 2.407064, label="0% population growth", srt=(-14.5))
	text(43.22775, 3.401599, label="1%", srt=(-14.5))
	text(48.51021, 4.622958, label="2%", srt=(-16))
	text(55.67089, 6.245621, label="3%", srt=(-18))
	text(64.35761, 8.321932, label="4%", srt=(-20))


	text(26.19916, 2.804078, label="hunter-\n gatherers", col="blue")

	text(30.95021, 8.619962, label="horticulturalists", col="orange")

	axis(2, at=seq(1,9, by=1), las=1)
	axis(1, at=seq(20,90, by=10), las=1)

	text(51.207, 8.18, label="1955", col=col_1955, cex=3.5)


	text(80.104, 3.63, label="2005", col=col_2005, cex=3.5)

#	text(56.78387, 3.349255, "world nations,\nyear 2005", font=2)

dev.off()

# reduce the number of nations, esp small ones that are probably throwing off
# maybe 


```



```{r, fig.height=15/2.54, fig.width=21/2.54}

d$my_col <- data_gradient(d$x, my.start=0, my.stop=3, colors=orangered )

tfr <- seq(1, 9, by=0.1)
le <- seq(20, 85, by=0.1)
background <- expand.grid(tfr, le)
colnames(background) <- c("tfr", "le")

background$ln_tfr <- log(background$tfr)
background$ln_le <- log(background$le)

background$x <- coef(m1)[1] + coef(m1)[2]*background$ln_tfr + coef(m1)[3]*background$ln_le

background$my_col <- data_gradient(background$x, my.start=0, my.stop=3, colors=orangered )


pdf('./pdf/test3.pdf', height=15/2.54, width=21/2.54)

	tar <- 1:nrow(d)

	plot(d$le, d$tfr, col=NA, frame.plot=FALSE, 
		ylim=c(1,9), xlim=c(20,80), 
		xlab="Life expectancy at birth", 
		ylab="Number of children per woman", las=1, axes=FALSE)

	points(background$le, background$tfr, col=background$my_col, pch=15)

	# curve(growth_isocline(x, -1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	
	foragers <- read.csv("./csv/forager-polygon.csv", stringsAsFactors=FALSE)
	pastoralists <- read.csv("./csv/pastoralist-polygon.csv", stringsAsFactors=FALSE)
	hort <- read.csv("./csv/horticulturalist-polygon.csv", stringsAsFactors=FALSE)

	tar <- 1:nrow(d)
	points(jitter(d$le[tar], amount=0.1), jitter(d$tfr[tar], amount=0.1), pch=16, col=col_alpha("gray", 0.3))


	polygon(hort$x, hort$y, border=NA, col=col_alpha("orange", 0.6) )
	polygon(foragers$x, foragers$y, border=NA, col=col_alpha("blue", 0.6) )
	polygon(pastoralists$x, pastoralists$y, border=NA, col=col_alpha("red", 0.6) )


	tar <- which(d$Year == 2000)
	points(d$le[tar], d$tfr[tar], pch=20)

	curve(growth_isocline(x, 0), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 2), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 3), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
	curve(growth_isocline(x, 4), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)



	# text

	# text(34.65842, 1.709144, label="-1%", srt=(-14.5))
	text(39.70611, 2.407064, label="0% population growth", srt=(-14.5))
	text(43.22775, 3.401599, label="1%", srt=(-14.5))
	text(48.51021, 4.622958, label="2%", srt=(-16))
	text(55.67089, 6.245621, label="3%", srt=(-18))
	text(64.35761, 8.321932, label="4%", srt=(-20))


	text(26.19916, 2.804078, label="hunter-\n gatherers", col="blue")

	text(30.95021, 8.619962, label="horticulturalists", col="orange")

	axis(2, at=seq(1,9, by=1), las=1)
	axis(1, at=seq(20,90, by=10), las=1)


	text(56.78387, 3.349255, "world nations,\nyear 2000", font=2)

dev.off()

# reduce the number of nations, esp small ones that are probably throwing off
# maybe 


```






Gurven and Kaplan report average life expectancy of:

- 29.23 for hunter/gatherers
- 38.80 for Horticulturalists
- 42.78 for pastoralists

They also report the following TFRs

- 5.7 for hunter-gatherers
- 6.2 for hort
- 5.6 for pastoralists

Each of these is a coordinate on the plot.

Before 1850, no human population existed. 

Between 1900 and 1955, set of human populations slide down. Followed by an explosion of growth, a second set of populations.
Arriving the sane place.

```{r}



```


```{r}

## animation

# for(i in 1900:2005){

# 	tar <- which(d$Year == i)
# 	#tar <- 1:nrow(d)

# 	plot(d$le[tar], d$tfr[tar], col=col_alpha(d$my_col[tar], 0.9), pch=20, frame.plot=FALSE, ylim=c(1,9), xlim=c(20,80), main=i, 
# 		xlab="Life expectancy at birth", 
# 		ylab="Number of children per woman (avg.)")
# 	curve(growth_isocline(x, -1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
# 	curve(growth_isocline(x, 0), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
# 	curve(growth_isocline(x, 1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
# 	curve(growth_isocline(x, 2), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
# 	curve(growth_isocline(x, 3), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
# 	curve(growth_isocline(x, 4), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)

# 	points(29.23, 5.7) # foragers average
# 	points(38.80, 6.2) # horticulturalists average
# 	points(42.78, 5.6) # pastoralists average

# 	Sys.sleep(0.3)

# }

```



```{r}


par(mfrow=c(2,2))


tar <- which(d$Year == 1850)

plot(d$le[tar], d$tfr[tar], col=col_alpha(d$my_col[tar], 0.9), pch=20, frame.plot=FALSE, ylim=c(1,9), xlim=c(20,80), main=1850, 
	xlab="Life expectancy at birth", 
	ylab="Number of children per woman (avg.)")
curve(growth_isocline(x, -1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 0), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 2), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 3), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 4), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)

points(29.23, 5.7, pch=20) # foragers average
points(38.80, 6.2, pch=20) # horticulturalists average
points(42.78, 5.6, pch=20) # pastoralists average



tar <- which(d$Year == 1900)

plot(d$le[tar], d$tfr[tar], col=col_alpha(d$my_col[tar], 0.9), pch=20, frame.plot=FALSE, ylim=c(1,9), xlim=c(20,80), main=1900, 
	xlab="Life expectancy at birth", 
	ylab="Number of children per woman (avg.)")
curve(growth_isocline(x, -1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 0), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 2), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 3), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 4), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)


points(29.23, 5.7, pch=20) # foragers average
points(38.80, 6.2, pch=20) # horticulturalists average
points(42.78, 5.6, pch=20) # pastoralists average



tar <- which(d$Year == 1955)

plot(d$le[tar], d$tfr[tar], col=col_alpha(d$my_col[tar], 0.9), pch=20, frame.plot=FALSE, ylim=c(1,9), xlim=c(20,80), main=1955, 
	xlab="Life expectancy at birth", 
	ylab="Number of children per woman (avg.)")
curve(growth_isocline(x, -1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 0), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 2), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 3), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 4), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)

points(29.23, 5.7, pch=20) # foragers average
points(38.80, 6.2, pch=20) # horticulturalists average
points(42.78, 5.6, pch=20) # pastoralists average




tar <- which(d$Year == 2000)

plot(d$le[tar], d$tfr[tar], col=col_alpha(d$my_col[tar], 0.9), pch=20, frame.plot=FALSE, ylim=c(1,9), xlim=c(20,80), main=2000, 
	xlab="Life expectancy at birth", 
	ylab="Number of children per woman (avg.)")
curve(growth_isocline(x, -1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 0), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 1), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 2), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 3), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)
curve(growth_isocline(x, 4), from=0, to=100, ylim=c(0, 8), add=TRUE, lty=2)

points(29.23, 5.7, pch=20) # foragers average
points(38.80, 6.2, pch=20) # horticulturalists average
points(42.78, 5.6, pch=20) # pastoralists average

# even if i dont have each data point, i can create a 2-D cloud based on the invidiaul observations of TFR and of life expectancy in each figure...i can use this!

	# my_rows <- which(d$Country.code=="BRA")
	# max_growth <- max(d$x[my_rows], na.rm=TRUE)
	# my_col <- data_gradient(max_growth, my.start=0, my.stop=3, colors=rev(rev(orangered)) )
	# points(d$le[my_rows], d$tfr[my_rows], type="l", col=col.alpha(my_col, 0.9))

	# my_rows <- which(d$Country.code=="CHN")
	# max_growth <- max(d$x[my_rows], na.rm=TRUE)
	# my_col <- data_gradient(max_growth, my.start=0, my.stop=3, colors=rev(rev(orangered)) )
	# points(d$le[my_rows], d$tfr[my_rows], type="l", col=col.alpha(my_col, 0.9))

```


```{r}

	# my_rows <- which(d$Country.code=="JPN")
	# max_growth <- max(d$x[my_rows], na.rm=TRUE)
	# my_col <- data_gradient(max_growth, my.start=0, my.stop=3, colors=rev(rev(orangered)) )
	# points(d$le[my_rows], d$tfr[my_rows], type="l", col=col.alpha(my_col, 0.9))


	# my_rows <- which(d$Country.code=="DEU")
	# max_growth <- max(d$x[my_rows], na.rm=TRUE)
	# my_col <- data_gradient(max_growth, my.start=0, my.stop=3, colors=rev(rev(orangered)) )
	# points(d$le[my_rows], d$tfr[my_rows], type="l", col=col.alpha(my_col, 0.9))


	# my_rows <- which(d$Country.code=="FRA")
	# max_growth <- max(d$x[my_rows], na.rm=TRUE)
	# my_col <- data_gradient(max_growth, my.start=0, my.stop=3, colors=rev(rev(orangered)) )
	# points(d$le[my_rows], d$tfr[my_rows], type="l", col=col.alpha(my_col, 0.9))

	# my_rows <- which(d$Country.code=="AUS")
	# max_growth <- max(d$x[my_rows], na.rm=TRUE)
	# my_col <- data_gradient(max_growth, my.start=0, my.stop=3, colors=rev(rev(orangered)) )
	# points(d$le[my_rows], d$tfr[my_rows], type="l", col=col.alpha(my_col, 0.9))


	# my_rows <- which(d$Country.code=="CAN")
	# max_growth <- max(d$x[my_rows], na.rm=TRUE)
	# my_col <- data_gradient(max_growth, my.start=0, my.stop=3, colors=rev(rev(orangered)) )
	# points(d$le[my_rows], d$tfr[my_rows], type="l", col=col.alpha(my_col, 0.9))


	# my_rows <- which(d$Country.code=="USA")
	# max_growth <- max(d$x[my_rows], na.rm=TRUE)
	# my_col <- data_gradient(max_growth, my.start=0, my.stop=3, colors=rev(rev(orangered)) )
	# points(d$le[my_rows], d$tfr[my_rows], type="l", col=col.alpha(my_col, 0.9))



	# my_rows <- which(d$Country.code=="GBA")
	# max_growth <- max(d$x[my_rows], na.rm=TRUE)
	# my_col <- data_gradient(max_growth, my.start=0, my.stop=3, colors=rev(rev(orangered)) )
	# points(d$le[my_rows], d$tfr[my_rows], type="l", col=col.alpha(my_col, 0.9))


```


```{r}
# # now plot lines by country - horrible!

# codes <- unique(d$Country.code)
# codes <- codes[-which(codes=="")]

# plot(1,1, col=NULL, pch=20, frame.plot=FALSE, ylim=c(1,9), xlim=c(20,80), 
# 	xlab="Life expectancy at birth", 
# 	ylab="Number of children per woman (avg.)")


# for(i in 1:length(codes)){
# 	my_rows <- which(d$Country.code==codes[i])
# 	max_growth <- max(d$x[my_rows], na.rm=TRUE)

# 	my_col <- data_gradient(max_growth, my.start=0, my.stop=3, colors=rev(rev(orangered)) )

# 	points(d$le[my_rows], d$tfr[my_rows], type="l", col=col.alpha(my_col, 0.9))
# }

# plot(d$le[tar], d$tfr[tar], col=col_alpha(d$my_col[tar], 0.5), pch=20, frame.plot=FALSE, cex=1, ylim=c(1,9), xlim=c(20,80), 
# 	xlab="Life expectancy at birth", 
# 	ylab="Number of children per woman (avg.)")


# }

```
